Faithful Variable Screening for\\ High-Dimensional Convex Regression

### 2015-11-19 tex/scam/summary.tex
minor revisions

In Section~\ref{sec:finitesample} 
we analyze the finite sample variable selection consistency of AC/DC, without assuming that the true
regression function $f_0$ is additive. Our analysis first establishes
a sufficient deterministic condition for variable selection 
consistency, and then considers a stochastic setting.
Our proof technique decomposes the KKT conditions for the optimization
in a manner that is similar to the now standard \emph{primal-dual
  witness} method~\citep{wainwright2009sharp}.

>>>>

In Section~\ref{sec:finitesample} 
we analyze the finite sample variable selection consistency of AC/DC under the model
\[
y_i = f_0(X_i) + w_i \quad \textrm{for $i=1,...,n$, }
\]
without assuming that the true
regression function $f_0$ is additive. Our analysis first establishes
a sufficient deterministic condition for variable selection 
consistency, and then considers a stochastic setting.
Our proof technique decomposes the KKT conditions for the optimization
in a manner that is similar to the now standard \emph{primal-dual
  witness} method~\citep{wainwright2009sharp}.

### 2015-11-19 tex/scam/supplement.tex
minor revisions

Suppose for all $k\in S^c$ and for all $i=1,...,n$, $\lambda_n \geq \mathsf{range}_k | \frac{32}{n} \hat{r}^\tran \mathbf{1}_{\pi_k(i:n)}|$, and $\max_{i=1,...,n-1} \frac{X_{\pi_k(i+1)k} - X_{\pi_k(i)k}}{\mathsf{range}_k} \geq \frac{1}{16}$, and $\mathsf{range}_k \geq 1$.

@@

\begin{theorem}
\label{thm:convex_consistent}
Let $\tilde{\sigma} \equiv \max(\sigma, B)$, and let $\hat{f}$ be the
minimizer of the restricted regression with $\lambda \leq 768 s
\tilde{\sigma} \sqrt{ \frac{1}{n} \log^2 np}$.
Suppose $n \geq c_1 s \sqrt{sB}$.
Then with probability at least $1-\frac{C}{n}$,
\begin{align}
\|f_0 - \hat{f} \|_P^2 - \| f_0 - f^* \|_P^2 
&\leq c B^2 \tilde{\sigma} \sqrt{ \frac{s^5}{n^{4/5}} \log^2 Cnp},
\end{align}
where $c_1$ is an absolute constant and $c, C$ are constants possibly dependent on $b$.
\end{theorem}

@@

\textbf{Step 1.} To bound term 1, we start from the definition of
$\hat{g}_k$ and obtain
\begin{align*}
\| y - \hat{f} - \hat{g}_k \|_n^2 + \lambda_n \| \hat{g} \|_\infty &\leq
   \| y - \hat{f} - (g^*_k - \bar{g}^*_k) \|_n^2 + \lambda_n \| (g^*-\bar{g}^*_k) \|_\infty \\
\| y - \hat{f} - \hat{g}_k \|_n^2 &\leq \| y - \hat{f} - (g^*_k-\bar{g}^*_k) \|_n^2 + \lambda_n B \\[10pt]
\| f_0 - \hat{f} - \hat{g}_k + w\|_n^2 & \leq \| f_0 - \hat{f} - (g^*_k-\bar{g}^*_k) + w \|_n^2 
   +\lambda_n B \\
\| f_0 - \hat{f} - \hat{g}_k \|_n^2 - \|f_0 -\hat{f} - (g^*_k-\bar{g}^*_k) \|_n^2 &\leq
   2 \langle w, \hat{g}_k - (g^*_k-\bar{g}^*_k) \rangle_n + \lambda_n B.
\end{align*}
$\bar{g}^*_k = \frac{1}{n} \sum_{i=1}^n g^*_k(X_{ik})$; we subtract it from $g^*_k$ again so that $g^*_k$ is empirically mean-zero.

@@

% And, with probability at least $1-\delta$, for all $\psi_U, \psi_L$:
% \begin{align*}
% \frac{1}{n} \sum_{i=1}^n ( \psi_U^2(X_i) - \psi_L^2(X_i)) &\leq 
%          \frac{1}{n} \sum_{i=1}^n | h_U^2(X_i) - h_L^2(X_i)| \\
%   &\leq \frac{1}{n}\sum_{i=1}^n | h_U(X_i) - h_L(X_i)| \, |h_U(X_i) + h_L(X_i)| \\
%   &\leq 2sB(\epsilon + \epsilon_{n,\delta})
% \end{align*}
Now we can bound $\| h \|_n^2 - \| h \|_P^2$ as
\begin{align}
\frac{1}{n} \sum_{i=1}^n \psi_L(X_i)^2 - \E \psi_U(X)^2  \leq
    \| h \|_n^2 - \| h \|_P^2 \leq
  \frac{1}{n} \sum_{i=1}^n \psi_U(X_i)^2 - \E \psi_L(X)^2  \label{eqn:hpsi_bound}
\end{align}
Since 
$\psi_L(X_i)^2$ and $\psi_U(X_i)^2$ are bounded random variables with
upper bound $(2sB)^2$, Hoeffding's inequality and union bound give that,
with probability at least $1-\delta$,, for all $\psi_L$ (and likewise $\psi_U$)
\begin{align*}
\left| \frac{1}{n} \sum_{i=1}^n \psi_L(X_i)^2 - 
   \E \psi_L(X)^2 \right| &\leq (2sB)^2 \sqrt{ \frac{sK^{**} (4sBc_u)^{1/2}}{\epsilon^{1/2}2n} + \frac{\log\frac{2}{\delta}}{2n}} 
\end{align*}
To simplify the expression, we will suppose that $\frac{sK^{**} (4sBc_u)^{1/2}}{\epsilon^{1/2}} \geq 2$ and that $\log \frac{2}{\delta} \geq 2$. The second supposition holds by assumption in the theorem. Once we calculate the proper values of $\epsilon$, we will verify that these first supposition holds under the assumption of the theorem also. Under these two suppositions, we have
\begin{align}
\label{eqn:simplified_psi_concentration}
\left| \frac{1}{n} \sum_{i=1}^n \psi_L(X_i)^2 - 
   \E \psi_L(X)^2 \right| &\leq (2sB)^2 \sqrt{ \frac{ sK^{**} (sBc_u)^{1/2} \log \frac{2}{\delta}}{ \epsilon^{1/2} n} }
\end{align}

@@

Since the probability measure $P$ is absolutely continuous with respect to the Lebesgue measure on $[0, 1]^p$, $\lambda (A_j) > 0$. Let $\lambda_1$ be the one dimensional Lebesgue measure on $[0,1]$ and let $A^1_j = \{ x_j \in [0,1] \,:\, \phi_j(x_j) > 0\}$. From the fact that $\lambda(A_j) > 0$, and that $A_j = A^1_j \times [0,1]^{p-1}$, $\lambda_1(A^1_j) > 0$. Same reasoning show that $\lambda_1(B^1_{j'}) > 0$ where $B^1_{j'}$ is similarly defined.\\

@@

>>>>

Suppose for all $k\in S^c$ and for all $i=1,...,n$, $\lambda_n \geq \mathsf{range}_k | \frac{32}{n} \hat{r}^\tran \mathbf{1}_{\pi_k(i:n)}|$, and $\max_{i=1,...,n-1} \frac{X_{\pi_k(i+1)k} - X_{\pi_k(i)k}}{\mathsf{range}_k} \leq \frac{1}{16}$, and $\mathsf{range}_k \geq 1$.

@@

\begin{theorem}
\label{thm:convex_consistent}
Let $\tilde{\sigma} \equiv \max(\sigma, B)$, and let $\hat{f}$ be the
minimizer of the restricted regression with $\lambda \leq 768 s
\tilde{\sigma} \sqrt{ \frac{1}{n} \log^2 np}$.
Suppose $n \geq c_1 s \sqrt{sB}$.
Then with probability at least $1-\frac{C}{n}$,
\begin{align}
\|f_0 - \hat{f} \|_P^2 - \| f_0 - f^* \|_P^2 
&\leq c B^2 \tilde{\sigma} \sqrt{ \frac{s^5}{n^{4/5}} \log^2 Cnp},
\end{align}
where $c_1$ is an absolute constant and $c, C$ are constants dependent on $b$.
\end{theorem}

@@

\textbf{Step 1.} To bound term 1, we start from the definition of
$\hat{g}_k$ and obtain
\begin{align*}
\| y - \hat{f} - \hat{g}_k \|_n^2 + \lambda_n \| \hat{g}_k \|_\infty &\leq
   \| y - \hat{f} - (g^*_k - \bar{g}^*_k) \|_n^2 + \lambda_n \| (g^*-\bar{g}^*_k) \|_\infty \\
\| y - \hat{f} - \hat{g}_k \|_n^2 &\leq \| y - \hat{f} - (g^*_k-\bar{g}^*_k) \|_n^2 + \lambda_n B \\[10pt]
\| f_0 - \hat{f} - \hat{g}_k + w\|_n^2 & \leq \| f_0 - \hat{f} - (g^*_k-\bar{g}^*_k) + w \|_n^2 
   +\lambda_n B \\
\| f_0 - \hat{f} - \hat{g}_k \|_n^2 - \|f_0 -\hat{f} - (g^*_k-\bar{g}^*_k) \|_n^2 &\leq
   2 \langle w, \hat{g}_k - (g^*_k-\bar{g}^*_k) \rangle_n + \lambda_n B.
\end{align*}
$\bar{g}^*_k = \frac{1}{n} \sum_{i=1}^n g^*_k(X_{ik})$; we subtract it from $g^*_k$ again so that $g^*_k$ is empirically mean-zero.

@@

% And, with probability at least $1-\delta$, for all $\psi_U, \psi_L$:
% \begin{align*}
% \frac{1}{n} \sum_{i=1}^n ( \psi_U^2(X_i) - \psi_L^2(X_i)) &\leq 
%          \frac{1}{n} \sum_{i=1}^n | h_U^2(X_i) - h_L^2(X_i)| \\
%   &\leq \frac{1}{n}\sum_{i=1}^n | h_U(X_i) - h_L(X_i)| \, |h_U(X_i) + h_L(X_i)| \\
%   &\leq 2sB(\epsilon + \epsilon_{n,\delta})
% \end{align*}
Now we can bound $\| h \|_n^2 - \| h \|_P^2$ as
\begin{align}
\frac{1}{n} \sum_{i=1}^n \psi_L(X_i)^2 - \E \psi_U(X)^2  \leq
    \| h \|_n^2 - \| h \|_P^2 \leq
  \frac{1}{n} \sum_{i=1}^n \psi_U(X_i)^2 - \E \psi_L(X)^2  \label{eqn:hpsi_bound}
\end{align}
Since 
$\psi_L(X_i)^2$ and $\psi_U(X_i)^2$ are bounded random variables with
upper bound $(2sB)^2$, Hoeffding's inequality and union bound give that,
with probability at least $1-\delta$,, for all $\psi_L$ (and likewise $\psi_U$)
\begin{align*}
\left| \frac{1}{n} \sum_{i=1}^n \psi_L(X_i)^2 - 
   \E \psi_L(X)^2 \right| &\leq (2sB)^2 \sqrt{ \frac{sK^{**} (4sBc_u)^{1/2}}{\epsilon^{1/2}2n} + \frac{\log\frac{2}{\delta}}{2n}} 
\end{align*}
To simplify the expression, we will suppose that $\frac{sK^{**} (4sBc_u)^{1/2}}{\epsilon^{1/2}} \geq 2$ and that $\log \frac{2}{\delta} \geq 2$. The second supposition holds by assumption in the theorem. Once we calculate the proper values of $\epsilon$, we will verify that the first supposition holds under the assumption of the theorem also. Under these two suppositions, we have
\begin{align}
\label{eqn:simplified_psi_concentration}
\left| \frac{1}{n} \sum_{i=1}^n \psi_L(X_i)^2 - 
   \E \psi_L(X)^2 \right| &\leq (2sB)^2 \sqrt{ \frac{ sK^{**} (sBc_u)^{1/2} \log \frac{2}{\delta}}{ \epsilon^{1/2} n} }
\end{align}

@@

Since the probability measure $P$ is absolutely continuous with respect to the Lebesgue measure on $[0, 1]^p$, $\lambda (A_j) > 0$. Let $\lambda_1$ be the one dimensional Lebesgue measure on $[0,1]$ and let $A^1_j = \{ x_j \in [0,1] \,:\, \phi_j(x_j) > 0\}$. From the fact that $\lambda(A_j) > 0$, and that $A_j = A^1_j \times [0,1]^{p-1}$, $\lambda_1(A^1_j) > 0$. Same reasoning shows that $\lambda_1(B^1_{j'}) > 0$ where $B^1_{j'}$ is similarly defined.\\

@@

\section{Simplifications when $p(\mathbf{x})$ is a product density}
\label{sec:prod_density_simplification}

If we assume that $p(\mathbf{x})$, the density over the covariates, is a product density, then some of the results can be strengthened and some can be simplified. Note that product densities satisfy the boundary flatness condition (definition~\ref{defn:boundary-point}) trivially and thus are settings under which convex functions are additively faithful. We suppose without loss of generality that the true function $f_0$ is convex and has mean-zero.

First, we will show that the unconstrained additive projection is convex under a product density and thus, the DC stage of the algorithm is not necessary. Let $\{ f^*_k \}$ be the unconstrained additive projection components.
\begin{align*}
 f^*_k &= \E[ f_0(X) - \sum_{k' \neq k} f^*_{k'}(X_{k'}) \given x_k ] \quad 
\textrm{(by Lemma~\ref{lem:general_int_reduction})} \\
     &= \E[ f_0(X) \given x_k ] \quad \textrm{(by independence of $X_k, X_{k'}$)} \\
  &= \int f_0(\mathbf{x}) p(\mathbf{x}_{-k} \given x_k ) d \mathbf{x}_{-k} \\
  &= \int f_0(\mathbf{x}) p(\mathbf{x}_{-k}) d\mathbf{x}_{-k} \quad
   \textrm{(by independence again)}
\end{align*}

If we see $f_0(\mathbf{x})$ as a set of functions $f_0(x_k, \mathbf{x}_{-k})$ of just $x_k$ indexed by $\mathbf{x}_{-k}$, we see that $f^*_k(x_k)$ is a weighted average of $f_0(x_k, \mathbf{x}_{-k})$'s and is thus convex.

Therefore, the unconstrained additive projection and the convex additive projection yield the same solution. Thus, the convex additive projection is also faithful and no additional concave postprocessing is needed.

Second, we can simplify the definition of the signal strength ($\alpha_+, \alpha_-$ from equation~\ref{eqn:signal_level_defn}) under a product density. $\alpha_-$ is no longer needed at all since the DC stage is unnecessary. As for $\alpha_+$, we have the following proposition:

\begin{proposition}
Let $\alpha_+$ be defined as in equation~\ref{eqn:signal_level_defn}. If $p(\mathbf{x})$ is a product density, then $\alpha_+ = \min_{k \,:\, f^*_k \neq 0} \E f^*_k(X_k)^2$. 
\end{proposition}

For any convex additive function $f$, we have that
\[
\alpha_+ = \E (f_0(X) - f(X))^2 - \E (f_0(X) - f^*(X))^2 \geq \E (f(X) - f^*(X))^2
\]
because $f^*$ is the orthogonal projection of $f_0$ onto the set of convex additive functions, which is itself a convex set.

Next, suppose that $f$ is a convex additive function such that $\supp(f) \subsetneq \supp(f^*)$, i.e., $f_k \neq 0 \Rightarrow f^*_k \neq 0$ but there exists some $k$ such that $f^*_k \neq 0$ and $f_k = 0$. 
\begin{align*}
\E (f(X) - f^*(X))^2 & = \E \left( \sum_{k=1}^p f_k(X_k) - f^*_k(X_k) \right)^2 \\
   &= \sum_{k=1}^p \E (f_k(X) - f^*_k(X))^2  \quad \textrm{(since $X_k, \perp X_{k'}$ for $k \neq k'$)} \\
   &\geq \min_{k \,:\, f^*_k \neq 0} \E f^*_k(X_k)^2 
\end{align*}

Now, let $k^* = \arg\min_k \E f^*_k(X_k)^2$ and let $f(\mathbf{x}) = \sum_{k\neq k^*} f_k^*(x_k)$. Then,

\begin{align*}
& \E (f_0(X) - f(X))^2 - \E (f_0(X) - f^*(X))^2 \\
 &=
   \E \big(f_0(X) - f^*(X) + f^*_{k^*}(X_{k^*}) \big)^2 - \E( f_0(X) - f^*(X))^2 \\
 &= \E \big[ 2(f_0(X) - f^*(X)) f^*_{k^*} (X_{k^*}) \big] + \E f^*_{k^*}(X_{k^*})^2 \\
 &= \E\big[ \E[ 2(f_0(X) - f^*(X)) \given X_{k^*}] f^*_{k^*}(X_{k^*}) \big] + \E f^*_{k^*}(X_{k^*})^2 \\
 &= \E f^*_{k^*}(X_{k^*})^2 
\end{align*}

In the second to last equality, we used the fact that $\E[ f_0(X) - f^*(X) \given X_{k^*}] = 0$ since $f^*_{k^*}(X_{k^*}) = \E[ f_0(X) \given X_{k^*} ]$ under a product density.

Thus, we have shown that $\alpha_+ \geq \min_{k \,:\, f^*_k \neq 0} \E f_k^*(X_k)^2$ and that equality is attained with $f(\mathbf{x}) = \sum_{k\neq k^*} f_k^*(x_k)$.
